Improving content discovery through combining linked data and data mining techniques

### 2016-04-25 report/chapters/04-implementation.tex
First pass of chapter 4 done

\subsection{Entity Extraction}

@@

    spotlight_response = requests.post(
            'http://spotlight.sztaki.hu:2222/rest/annotate',
            data={'text': text, 'confidence': 0.8, 'support': 20},
            headers={'Accept': 'application/json'},
            timeout=600, proxies={'http': ''}
        )

    if spotlight_response.ok:
        r_json = spotlight_response.json()
        if 'Resources' in r_json:
            annotations = {
                resource['@URI']
                for resource in r_json['Resources']
                }
            if annotations:
                for entity in annotations:
                    g.add((URIRef(iri), FOAF.topic, URIRef(entity)))
    return g
  \end{lstlisting}
\end{centering}

@@

@@

@@

@@

@@

@@

>>>>

\subsection{Entity Extraction}
\label{sec:impl-entity-extraction}

@@

    spotlight_response = requests.post(
        'http://spotlight.sztaki.hu:2222/rest/annotate',
        data={
          'text': text, 'confidence': 0.8, 'support': 20
        },
        headers={'Accept': 'application/json'},
        timeout=600)

    if spotlight_response.ok:
        r_json = spotlight_response.json()
        if 'Resources' in r_json:
            annotations = {
                resource['@URI']
                for resource in r_json['Resources']
                }
            if annotations:
                for entity in annotations:
                    g.add((
                        URIRef(iri),
                        FOAF.topic,
                        URIRef(entity)))
    return g
  \end{lstlisting}
\end{centering}

@@

In listings~\ref{lst:extract-entities}, we see a simple Python
function that uses the Beautiful Soup
library\footnote{https://www.crummy.com/software/BeautifulSoup/} to
search across all links in a document and infer a ``related''
property on the assumption that pages like to other pages that are in
some way related.

\begin{centering}
  \begin{lstlisting}[
      label=lst:find-links,
      language=python,
      caption={Python function generates triples based on links to other pages}]
import requests
from rdflib import Graph, URIRef, Namespace

DBPROP = Namespace('http://dbpedia.org/property/')

def find_links(iri):
    g = Graph(identifier=iri)

    text = requests.get(iri).text
    doc = BeautifulSoup(text)

    # For all <a> tags that have an href attribute...
    for a_tag in doc.find_all('a', attrs={
        'href': re.compile('.+')
    }):
        # ... handle href values being relative links ...
        other_iri = urljoin(iri, a_tag.attrs['href']).strip()

        # ... infer triple if IRI is to another BBC page
        if bbc_url.match(url):
            g.add((
                URIRef(iri),
                URIRef(DBPROP.related),
                URIRef(other_iri),
            ))

    return g
  \end{lstlisting}
\end{centering}

This falls foul of similar issues to entity extraction as described
in section~\ref{sec:impl-entity-extraction} in that site-wide
navigation links and other supporting page elements may link to
very general pages (e.g. a persistent link to ``home'' on every page).

With appropriate feature selection, this may have no impact, but for
this experiment a simple heuristic was employed to narrow down to
the ``interesting'' part of BBC pages.

Another feature in listings~\ref{lst:find-links} is a restriction
to consider only links to other BBC pages. There may be a positive
or negative effect in addtionally noting external pages linked to from
content pages, but that is left to future research to consider.

@@

The combinatorial consequence of the usages and non-usages of the
three extraction techniques described so far is that the
experimental system developed is already comparing seven ($2^n$ less
the case where no extraction is employed).

In order to maintain focus in the experiment, only one of the
enrichment techniques described in chapter~\ref{design} was
implemented and its usage and non-usage merely doubled the
number of systems for comparison to fourteen.

This proved sufficient to give indicative results as to whether there
is value added through any enrichment, but future research could
certainly compare the merits of different enrichment approaches.

\begin{centering}
  \begin{lstlisting}[
      label=lst:enrich,
      language=python,
      caption={Python function that enriches a graph via deferencing objects}]
import rfc3987

def get_objects(graph):
    return {str(row.o) for row in graph.query('''
            SELECT DISTINCT ?o
            WHERE {
              ?iri ?p ?o .
            }
            ''', initBindings={'iri': graph.identifier})}

def enrich(graph):
    new_graph = copy(graph)

    for o in get_objects(graph):
        if rfc3987.match(o, rule='IRI'):
            new_graph.parse(iri2uri(o))

    return new_graph
  \end{lstlisting}
\end{centering}

Listings~\ref{lst:enrich} illustrates a simple function to
deference all objects of triples where the current IRI is a subject.
Note the convention to set the graph's own identifier to that of
the content item of interest.

Using SPARQL in this way opens up possibilities to perform multiple
queries to choose IRIs to dereference. For example, we may wish to
include semantics for objects reachable by following two predicates
on the RDF graph and consider much more indirect data. Evaluating
the utility of this is left for future work.

@@

The feature generation strategy employed was dervied from a distilled
portion of the approach introduced by Paulheim
and F\"urnkranz\cite{paulheim2012unsupervised}.

Listings~\ref{lst:generation} shows an illustrative, recursive
function that ``walks'' the RDF graph starting with the content
item's IRI as the initial subject.

\begin{centering}
  \begin{lstlisting}[
      label=lst:generation,
      language=python,
      caption={Python function that generates feature sets from RDF graphs}]
from rdflib Literal, URIRef

def generate(g, subject=None, depth=3, features=None, prefix=''):
    if depth <= 0:
        return
    if not features:
        features = {}
    if not subject:
        subject = g.identifier
    for p, o in g.predicate_objects(subject=subject):
        if type(o) == URIRef:
            new_prefix = prefix + clean(p) + '_'
            features[new_prefix + clean(o)] = True
            generate(g, depth=(depth - 1), features=features, prefix=new_prefix, subject=o)
        elif type(o) == Literal:
            features[clean(p)] = str(o)

    return features
  \end{lstlisting}
\end{centering}

The \texttt{clean} function simply provides some cosmetic cleaning
up of predicate IRIs to prefer CURIEs as they are more compact. A
typical example feature might then be
\texttt{foaf:topic\_dbpedia:Albert\_Einstein: true} where the leaf
object is an entity or
\texttt{ogp:title: "Gandhi: Reckless teenager to father of India"} if
the leaf object is a literal.

@@

A simple selection heuristic was employed, again inspired by the work
of Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}, to go
some way to reduce potential noise in the subsequent machine learning
phase.

In this experiment, all features that have the same value or entirely
unique values were removed and values that occurred twice or more
for a given feature had to make up over 5\% of possible values for
that feature. That is, if a feature had over 95\% of its values being
unique or all the same, then it was rejected.

@@

Hierarchical, agglomerative clustering was implemented by implementation
of a \texttt{cluster} function that was merge the two ``nearest''
clusters on each invocation. This assumes that all content items not
yet in a cluster are implicitly in a singleton cluster and the ability
to merge one at a time allowed for control of when to cease merging.

The closeness of two potential, candidate clusters was calculated via
a max linkage on the basis that it ensures that any two content items
within a cluster to be considered related. This avoids the so-called
\emph{chaining phenomenum} where single linkage may create clusters
where two unrelated items are members because they are each related
to an item of chain of related items
therebetween.\cite{everitt2011hierarchical}

The distance between content items was defined by the Jaccard distance
due to its suitability for binary and categorical data.

@@

With all of the components described so far in place, an experiment
was run processing nearly 10,000 content items from the BBC website
with each going into one of fourteen combination of extraction
and enrichment approaches.

For each approach, hiearchical clustering was applied repeatedly
until the next merge would produce a cluster with cohesion below 0.5.
This is reasonably arbitrary as each technique will have different
concepts of cohesion, but it provided a strong basis on which to
evaluate each approach on whether its cohesion meaningfully predicts
whether a human would agree that cluster contains related items.
